Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-17 nips_2018.tex
placeholder list tables

How should we get the data to train $q$?
This is nonobvious  because \system is only weakly supervised (i.e., learns only from tasks and not from (program, task) pairs).
One approach is to sample programs from the DSL,
run them to get their input/outputs,
and then train $q$ to predict the program from the input/outputs.
This approach is like how a Helmholtz machine
trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
The advantage of ``Helmholtz machine'' training is that
we can draw unlimited samples from the DSL,
training on a large amount of data.
Another approach is
 self-supervised learning,
training $q$ on the (program, task)
pairs discovered by the algorithm so far.
The advantage of self-supervised learning
is that the training data is much higher quality,
because we are training on the actual tasks.
Due to these complementary advantages,
we train on both these sources of data.

@@

We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting fragments of
programs already in the frontiers (Fig.~\ref{fragmentExample}).
An important point here is that we are \emph{not} simply adding
subexpressions of programs to $\mathcal{D}$.  Instead, we are
extracting fragments that unify with programs in the frontiers.  This
idea of storing and reusing fragments of expressions comes from
Fragment Grammars~\cite{tim} and Tree-Substitution
Grammars~\cite{cohn2010inducing}, and is closely related to the idea
of antiunification~\cite{henderson2013cumulative}.

@@

\begin{comment}
\begin{table}
\centering
\begin{tabular}{lll}
  \toprule
  Name & Input & Output \\\midrule
  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
  len & [3\, 5\, 12\, 1] & 4 \\
  has-2 & [4\, 5\, 7\, 4] & \code{false} \\
  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
  count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
  rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
  pow-3 & [10\, 4\, 13] & [1000\, 64\, 2197] \\
  keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
  \bottomrule
\end{tabular}
\caption{Sample tasks from our list function domain %% We used 15 input/output
  %% examples for each task to reduce ambiguity in the task's function,
  %% ensuring solutions achieve the desired concept.
}
\label{listexamples}
\end{table}

\begin{table}[t]
  \centering
  \begin{tabular}{l}
    \toprule
    $f_0($\code{i$,\ell$}$) \,=\, $\code{(singleton (index i $\ell$))}\\
    \hspace{0.5cm}($f_0$: \emph{Put the $i$-th index into a list})\\
    $f_1($\code{i$,\ell$}$) \,=\, $\code{(++ $\ell$ ($f_0$ i $\ell$))}\\
    \hspace{0.5cm}($f_1$: \emph{Append the $i$-th index})\\
    $f_2($\code{n$,\ell$}$) \,=\, $\code{(any ($\lambda$ (x) (= n x)) $\ell$)}\\
    \hspace{0.5cm}($f_1$: \emph{Whether $n$ appears in the list})\\
    $f_3($\code{i$,\ell$}$) \,=\, $\code{(index (negate i) (sort $\ell$))}\\
    \hspace{0.5cm}($f_3$: \emph{Get the $i$-th largest number})\\
    $f_4($\code{n$,\ell$}$) \,=\, $\code{(mapi ($\lambda$ (i x) (mod x n)) $\ell$)}\\
    $f_5($\code{k$,\ell$}$) \,=\, $\code{(mapi ($\lambda$ (i x) (+ x k)) $\ell$)}\\
    $f_6($\code{k$,$n$,\ell$}$) \,=\, $\code{($f_4$ n ($f_5$ k $\ell$)}\\
    \hspace{0.5cm}($f_6$: \emph{Caesar shift of $k$ in integers modulo $n$})\\
  \bottomrule
  \end{tabular}
  \caption{Some learned DSL primitives for list functions}\label{listinduced}
\end{table}

  \end{comment}

@@

We automatically generated 109 text editing tasks  %% with 4 input/output examples each
(Fig.~\ref{exampleTextProblem}).
At first, \system cannot find any correct programs for most of the tasks.
After three iterations, it assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

@@

\begin{table}[t]
  \centering
    \begin{tabular}{l}
      \toprule
      $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
        %($\lambda$ ($\lambda$ (foldr $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
  \\\bottomrule
    \end{tabular}
  \caption{Some text editing learned subroutines }\label{textPrimitives}
  \end{table}

@@

\begin{wraptable}{r}{0.6\textwidth}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{llllll}
  \toprule& Ours& \begin{tabular}{c}
    Ours\\(no NN)
    \end{tabular}& RF/DC & PCFG & Enum

@@

 \noindent \emph{Deep learning:} Much recent work in the ML community has
 focused on creating neural networks that regress from
 input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. %This family of work is closest to our RF/DC baseline.
 These neural networks are typically trained with strong supervision (i.e., with annotated ground-truth programs) on massive data sets (i.e., hundreds of millions~\cite{devlin2017robustfill}).
 Our work  considers a weakly-supervised regime where ground truth programs are not provided and
 the agent must learn from a few hundred tasks.

 \noindent \emph{Inventing new subroutines for program induction:}
 Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting. We find this work inspiring and motivating,
 and extend it along two dimensions: (1) we propose a new algorithm for
 inducing reusable subroutines, based on Fragment Grammars~\cite{tim};
 and (2) we show how to combine these techniques with bottom-up neural recognition models.
 Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in ILP~\cite{DBLP:conf/ecai/LinDETM14}.

 Our work is an instance of
 Bayesian Program
 Learning (BPL; see~\citep{lake2015human,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}). Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),
 and our contribution here is a general way of doing BPL with less hand-engineering of the DSL.

@@

We contribute an algorithm, \systemEnding, that learns to program by bootstrapping a DSL with new
domain-specific primitives that the algorithm itself discovers,
together with a neural recognition model that learns how to
efficiently deploy the DSL on new tasks. We believe this integration
of top-down symbolic representations and bottom-up neural networks --
both of them learned -- could help make program induction systems more
generally useful for AI. Many directions remain open. Two immediate
goals 
are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in
some domains over the generic methods used here.
Another direction is to
explore DSL meta-learning: Can we find a \emph{single} universal
primitive set that could effectively bootstrap DSLs for new domains,
including the four domains considered,  but also many others?

>>>>

How should we get the data to train $q$?
This is nonobvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from (program, task) pairs).
One approach is to sample programs from the DSL,
run them to get their input/outputs,
and then train $q$ to predict the program from the input/outputs.
This is like how a Helmholtz machine
trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
The advantage of ``Helmholtz machine'' training is that
we can draw unlimited samples from the DSL,
training on a large amount of data.
Another approach is
 self-supervised learning,
training $q$ on the (program, task)
pairs discovered by search.
The advantage of self-supervised learning
is that the training data is much higher quality,
because we are training on the actual tasks.
Due to these complementary advantages,
we train on both these sources of data.

@@

We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting fragments of
programs already in the frontiers (Fig.~\ref{fragmentExample}).
An important point here is that we are \emph{not} simply adding
subexpressions of programs to $\mathcal{D}$, as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316} and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
extracting fragments that unify with programs in the frontiers.  This
idea of storing and reusing fragments of expressions comes from
Fragment Grammars~\cite{tim} and Tree-Substitution
Grammars~\cite{cohn2010inducing}, and is closely related to the idea
of antiunification~\cite{henderson2013cumulative}.

@@

\begin{figure*}
\makebox[\textwidth][c]{
\begin{minipage}{0.6\textwidth}
  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    has-2 & [4\, 5\, 7\, 4] & \code{false} \\
    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    \bottomrule
  \end{tabular}
  \captionof{table}{Sample tasks from our list function domain}
  \label{listexamples}
\end{minipage}
\hfill
\begin{minipage}{0.6\textwidth}
  \centering
  \begin{tabular}{l}
    \toprule
    $f_0($\code{i$,\ell$}$) \,=\, $\code{(TODO)}\\
    \hspace{0.5cm}($f_0$: \emph{TODO})\\
  \bottomrule
  \end{tabular}
  \captionof{table}{Some learned DSL primitives for list functions}\label{listinduced}
\end{minipage}
  }
\end{figure*}

@@

We automatically generated 109 text editing tasks with 4 input/output examples each
%(Fig.~\ref{exampleTextProblem}).
At first, \system cannot find any correct programs for most of the tasks.
After three iterations, it assembles a DSL (Fig.~\ref{firstPageFigure} \& center of Fig.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

@@

%% \begin{table}[t]
%%   \centering
%%     \begin{tabular}{l}
%%       \toprule
%%       $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
%%       \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
%%       $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
%%         \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
%%       $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
%%         \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
%%         $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (if (= c x) nil (cons x a))))}\\
%%         \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
%%         %($\lambda$ ($\lambda$ (foldr $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
%%   \\\bottomrule
%%     \end{tabular}
%%   \caption{Some text editing learned subroutines }\label{textPrimitives}
%%   \end{table}

@@

\begin{wraptable}{r}{0.6\textwidth}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{llllll}
  \toprule& Ours& \multicolumn{1}{c}{\begin{tabular}{c}
    Ours\\(no NN)
    \end{tabular}}& RF/DC & PCFG & Enum

@@

 \noindent \textbf{Deep learning:} Much recent work in the ML community has
 focused on creating neural networks that regress from
 input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. %This family of work is closest to our RF/DC baseline.
\systemEnding's recognition model draws heavily from this line of work, particularly from~\cite{menon2013machine}.
We see these prior works as operating in a different regime: typically, they train with strong supervision (i.e., with annotated ground-truth programs) on massive data sets (i.e., hundreds of millions~\cite{devlin2017robustfill}).
 Our work   considers a weakly-supervised regime where ground truth programs are not provided and
 the agent must learn from at most a few hundred tasks,
 which is facilitated by our ``Helmholtz machine'' style recognition model training (Sec.~\ref{recognitionSection}).

 \noindent \textbf{Inventing new subroutines for program induction:}
 Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting. We find this work inspiring and motivating,
 and extend it along two dimensions: (1) we propose a new algorithm for
 inducing reusable subroutines, based on Fragment Grammars~\cite{tim};
 and (2) we show how to combine these techniques with bottom-up neural recognition models.
 Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.

\noindent\textbf{Bayesian Program
 Learning:} Our work is an instance of
 Bayesian Program
 Learning (BPL; see~\citep{lake2015human,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}). Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),
 and our contribution here is a general way of doing BPL with less hand-engineering of the DSL.

@@

We contribute an algorithm, \systemEnding, that learns to program by bootstrapping a DSL with new
domain-specific primitives that the algorithm itself discovers,
together with a neural recognition model that learns how to
efficiently deploy the DSL on new tasks. We believe this integration
of top-down symbolic representations and bottom-up neural networks --
both of them learned -- could help make program induction systems more
generally useful for AI. Many directions remain open. Two immediate
goals 
are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in
some domains over the generic methods used here.
Another direction is to
explore DSL meta-learning: Can we find a \emph{single} universal
primitive set that could effectively bootstrap DSLs for new domains,
including the three domains considered,  but also many others?
